Transferability of clinical prediction models for early trauma care

### 2019-09-24 ManuscriptMarkdown.Rmd
Update ManuscriptMarkdown.Rmd

---
title: "Transferability of clinical prediction models for early trauma care"
author: 
- Martin Henriksson
- Dell D Saulnier
- Johanna Berg
- Martin Gerdin Wärnberg
date: "September 18, 2019"
documentclass: "article"
output:
  word_document:
    reference_docx: transferstyle.docx
header-includes:
   - \usepackage{hyperref}
bibliography: library.bib
csl: bmcemerg.csl
---
```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)

@@

  **Background:** Trauma results in approximately 3.8 million deaths annually. Clinical
  prediction models are often used for triage in early trauma care. Previous research has
  indicated that transferring these models between different settings might adversely affect 
  model performance.

@@

  **Results:** A total of `r total.number.of.patients ` patients were included. The model
  validation mistriage rate ranged from `r lowest.val.mistriage` 
  to `r highest.val.mistriage`, and the transferred model
  mistriage rate ranged from `r lowest.tran.mistriage` to `r highest.tran.mistriage` 
  across all samples. Performance after transfer when compared to the receiving sample validation 
  performance resulted in substantial uncertainty in all samples. The worst performance after 
  transfer was found in the `r colnames(highest.comp.mistriage)` 
  sample, with a mistriage difference of `r highest.comp.mistriage`.

  **Conclusions:** In all samples, model transfer was associated with some degree of uncertainty. 
  Models developed in samples with a high number of patients performed the best after transfer. Affirming 
  this, the worst performance after transfer was the low-volume model transferred to the high-volume 
  sample, resulting in a significant mistriage increase of `r highest.comp.mistriage`, this was
  primarily due to increased overtriage.

@@

# Introduction
Trauma, defined as physical injuries to a host by outside objects [@RN63], 
accounts for approximately 458.5 million cases annually across the globe [@RN999] 
and an estimated 3.8 million deaths [@RN1; @RN21]. Each year, 9% of global 
deaths are the result of trauma, with the leading causes being road traffic accidents, 
suicide, and homicide. Predictions indicate that the incidence of these causes is 
likely to increase by 2030 [@RN42].

>>>>

---
title: "Transferability of clinical prediction models for early trauma care"
author: 
- Martin Henriksson
- Dell D Saulnier
- Johanna Berg
- Martin Gerdin Wärnberg
date: "September 24, 2019"
documentclass: "article"
output:
  word_document:
    reference_docx: transferstyle.docx
header-includes:
   - \usepackage{hyperref}
bibliography: library.bib
csl: bmcemerg.csl
---
```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = TRUE)

@@

  **Background:** Trauma results in approximately four million deaths annually. Clinical
  prediction models are often used for triage in early trauma care. Previous research has
  indicated that transferring these models between different settings might adversely affect 
  model performance.

@@

  **Results:** A total of `r total.number.of.patients ` patients were included. Model
  validation mistriage rate ranged from `r lowest.val.mistriage` 
  to `r highest.val.mistriage`, and the transferred model
  mistriage rate ranged from `r lowest.tran.mistriage` to `r highest.tran.mistriage`, 
  across all samples. The worst performance after transfer was the low-volume model transferred 
  to the high-volume sample, resulting in a significant mistriage rate increase of `r highest.comp.mistriage`, 
  primarily due to increased overtriage.

  **Conclusions:** In all samples, model transfer was associated with some degree of uncertainty. 
  Models developed in samples with high numbers of patients performed best after transfer, and the
  model developed in the low-volume sample performed worst after transfer.

@@

# Introduction
Trauma, defined as physical injuries to a host by outside objects [@RN63], 
accounts for approximately 458.5 million cases annually across the globe [@RN999] 
and around four million deaths [@RN1; @RN21]. Each year, 9% of global 
deaths are the result of trauma, with the leading causes being road traffic accidents, 
suicide, and homicide. Predictions indicate that the incidence of these causes is 
likely to increase by 2030 [@RN42].
